ms_2_results

### 2018-02-12 man/ms_2_results.Rmd
add review comments of GGI

## Vegetation Greenness

Trends analysis showed that 78.9 % of the pixels of *Quercus pyrenaica* forests experienced an EVI positive trend for the 2000-2016 period, of which 31.67 % were significant trends. The strongest trends were observed in southwestern populations (Figure S2). Vegetation greenness of *Q. pyrenaica* forests were lower during both the 2005 and 2012 years than the greenness observed for the reference period (Figure 2a). The lowest values for EVI standardized anomalies were recorded in 2005 being singnificantly lower (-2.285 ± 0.029) than 2012 (-0.418 ± 0.029) (LSMEANS, t.ratio = -45.358; p_value < 0.0001), particularly for northern populations (Figure 2b). Reduction in annual EVI mean was considerably higher in northern populations than in southern ones during the 2005 drought (Figure 2b).

According to the standardized anomalies, *Q. pyrenaica* forests suffered a browning during 2005 drought event (99.36 % and 79.37 % of the pixels for northern and southern populations respectively), however no changes were observed for the 2012 drought (Figure S3).

@@

## Responses to drougth: greenness

*Q. pyrenaica* forest showed significantly lower resistance to 2005 drought event than to 2012 one (Table 3; Figure 3a). The 2005 drought reduced the greenness of oak to 85.8 % while the 2012 reduced 94.3 % (Table S1). Southern populations showed significantly higher values of resistance to drought than northern ones (Table 3), except for 2012 where non-significant differences were recorded (Table S1. Figures 3a, 4a).

@@

The oak forests recovered their greenness significantly more rapidly after the 2005 drought than after 2012 (Tables 3, S1). In the period after 2005 drought, greeness achieved was 112 % (Rc = 1.12) and after 2012 was 105.7 % (Rc = 1.057) (Table S1). A similar recovery after the 2005 and 2012 drought event was observed for southern populations (p = 0.2453; Figure 4a), whilst the northern populations showed a significantly greater recovery after the 2005 drought than after the 2012 drought (Figure 4a).

@@

Resilience values were significantly higher for the 2012 drought event than for 2005 (Tables 3, S1), although both values were close to 1 indicating that greenness level was rather similar after each disturbance event (Table S1). The southern populations showed higher resilience values than the northern ones, although they were not significantly different for 2005 drought (p = 0.036; Figures 3a, 4a).

@@

## Growth trends
The trees of the southern population were older and larger than northern ones, particularly for the high elevation site (Ca-High) which had bigger and taller trees than the other sites (Tables 1, 2). Competence indices were similar among sites (Table 1). The basal area increment (BAI) were different between sites (Figure 5). El sitio mas elevado (Ca-High) mostró mayores valores de BAI que los otros sitios, mostrando además una tendencia positiva a partir de 1970. Se observan diferencias en cuanto al crecimiento entre los sitios de las poblaciones del sur. Ca_Low se parece en crecimiento mas a SJ que a Ca_High.

@@

Sorprende que en los sitios del sur (caH y caL), en los últimos años ha habido crecimientos muy grandes. Por ejemplo, en 2010 se observó uno de los mayores valores de RWI de toda la cronología tanto en caH como en caL. Algo similar ocurrió en sj para los años 2003 y 2013.

## Responses to drougth: tree-growth

Although no significant differences were observed in the resistance of oak growth between the two drought events (Table 3), the 2005 drought reduced growth more than that of 2012 (Rt = 0.721 and 0.819 respectively) (Table S2). The northern site, which is under a drier cliamte, showed resistance values lower than those of the southern site, especially for the 2005 drought event where the growth was reduced to 44.5 % respect to that of the preceding period (Figures 3b, 4b).

@@

Oak forests recovered their growth more rapidly after 2012 than after 2005 staying at levels of Rc < 1 (Table S2). Northern populations showed signigicantly higher values of recovery than southern ones (Tables 3 and S2), which even not reach recovery values of Rc = 1 (Figure 3b, 4b).

@@

## Climate and tree-growth
Tree-growth exhibited a significant greater response to water availability more than to temperatures, as it can be expected for a Mediterranean site and a species located closer to its xeric/dry distribution limit (rear-edge). Cumulative precipitation of the hydrological year was the climatic variable with a higher relationship with growth and influenced positively the tree-growth for the both northern and southern populations (Figure 7a). Precipitation of previous december was also positively correlated with tree growth in the northern population and in the highest location of the southern population. Hydrological, Spring and Summer SPEI showed a strong positive correlation with tree-growth (Figure 7b), specially for northern population (r > 0.6), which can be interpreted as higher sensitivity to drought of a drier site [@GeaIzquierdo2014]. Temperatures were much less related to growth than moisture variables. Yet spring maximum temperature was the most significant limitation for the tree-growth of the southern population at both elevations while maximum temperatures of the previous August were the only negative significant for the northern one (Figures 7c and 7d). Surprisingly minimum and maximum temperatures of current September positively influenced the tree-growth in the northern population.

## Disturbances
The analysis of growth changes revealed differences between sites (Figures 8 and 9). En el sitio del norte osbervamos dos eventos de liberación importantes (GC > 50 %), que ocurren en mas de la mitad de los árboles muestreados (Figures 8 and 9). El primero de ellos desde 1940 hasta 1950, mientras que el segundo entre 1994 y 2001. Estos periodos se relacionan bien con eventos antrópicos: mineria (el primero de ellos) y actividades forestales el segundo (esta frase de la minería y demás la dejamos para la discusión??). Estos periodos se alternan con periodos de supresión. En los sitios del sur (caH y caL), sin embargo se observan algunos episodios débiles de supresión, aunque estos no se observan para los últimos años, de hecho incluso se observa una débil liberación (Figure 8) afectando a mas de la mitad de los árboles muestreados (Figure 9).

>>>>

## Time trends in vegetation greenness

The analysis of greenness time trends showed that 78.9 % of the pixels of *Quercus pyrenaica* forests experienced an EVI positive trend for the 2000-2016 period, of which 31.67 % were significant trends. The strongest trends were observed in southwestern populations (Figure S2), in which the Cañar site are included Vegetation greenness of *Q. pyrenaica* forests were lower during both the 2005 and 2012 years than the greenness observed for the reference period (Figure 2a). The lowest values for EVI standardized anomalies were recorded in 2005 being singnificantly lower (-2.285 ± 0.029) than those in 2012 (-0.418 ± 0.029) (LSMEANS, t.ratio = -45.358; p_value < 0.0001), particularly for northern populations (Figure 2b). Reduction in annual EVI mean was considerably higher in northern populations than in southern ones during the 2005 drought (Figure 2b).

According to the standardized anomalies, *Q. pyrenaica* forests suffered a "browning episode" during 2005 drought event (99.36 % and 79.37 % of the pixels for northern and southern populations respectively), yet no changes in greenness were observed in response to the 2012 drought (Figure S3).

@@

## Greenness resilience to drought events

*Q. pyrenaica* forest showed significantly lower resistance (*Rt*) to the 2005 drought event than to that in 2012 (Table 3; Figure 3a). The 2005 drought reduced the greenness of oaks to a 85.8 % respect the reference period while the 2012 reduced it 94.3 % (Table S1). Southern populations showed significantly higher values of resistance to drought than northern ones (Table 3), except for 2012 where non-significant differences were recorded (Table S1. Figures 3a, 4a).

@@

The oak forests recovered (*Rc*) their greenness significantly more rapidly after the 2005 drought than after 2012 (Tables 3, S1). In the 3-year period after the 2005 drought, greenness achieved was 112 % (Rc = 1.12) and after 2012 was 105.7 % (Rc = 1.057) (Table S1). A similar recovery after the 2005 and 2012 drought event was observed for southern populations (p = 0.2453; Figure 4a), whilst the northern populations showed a significantly greater recovery after the 2005 drought than after the 2012 drought (Figure 4a).

@@

Resilience (*Rs*) was significantly higher for the 2012 drought event than for 2005 (Tables 3, S1), although both values were close to 1 indicating that greenness level was rather similar after each disturbance event (Table S1). The southern populations showed higher resilience values than the northern ones, although they were not significantly different for 2005 drought (p = 0.036; Figures 3a, 4a).

@@

## Radial growth trends and climate

Likely as a consequence of their different management origin (i.e. land-use legacy), the trees of the southern population were older than northern ones, particularly for the high elevation site (Ca-High) which had bigger and taller trees than the other sites (Tables 1, 2). The competition index was similar among sites but plot basal area was greatest in Ca-High (Table 1). Yet the basal area increment (BAI) growth in the highest plot in altitude (Ca-High) was the greatest  (Figure 5). Trees from this location showed in addition a positive growth trend since the late 1970s. This positive trend was not found in any of the other two locations and the two altitudes sampled in the Southern populations showed differences in growth, being more similar Ca_Low with growth in the Northern population (SJ) than with Ca_High (Figure 5). 
`VAS POR AQUI CARPINTO (BREAK)`

@@

Sorprende que en los sitios del sur (caH y caL), en los últimos años ha habido crecimientos muy grandes. Por ejemplo, en 2010 se observó uno de los mayores valores de RWI de toda la cronología tanto en caH como en caL. Algo similar ocurrió en sj para los años 2003 y 2013. 
")
```

## Resilience of radial tree-growth to drougth events

Although no significant differences were observed in the resistance (*Rt*) of oak radial growth between the two drought events (Table 3), the 2005 drought reduced growth more than that of 2012 (Rt = 0.721 and 0.819 respectively) (Table S2). Similar to results for greenness, the northern site, which is under a drier climate, showed resistance values lower than those of the southern site, especially for the 2005 drought event where the growth was reduced to 44.5 % respect to that of the preceding period (Figures 3b, 4b).

@@

Oak forests recovered (*Rc*) their growth more rapidly after 2012 than after 2005 staying at levels of Rc < 1 (Table S2). Northern populations showed signigicantly higher values of recovery than southern ones (Tables 3 and S2), which even not reach recovery values of Rc = 1 (Figure 3b, 4b).

@@

## Climate and tree-growth
Tree-growth exhibited a significant greater response to water availability than to temperatures, as it can be expected for a Mediterranean site and a species located closer to its xeric/dry distribution limit (rear-edge). Cumulative precipitation of the hydrological year was the climatic variable with the highest relationship with growth and influenced similarly and positively the tree-growth for both northern and southern populations (Figure 7a). Precipitation of previous december was also positively correlated with tree growth in the northern population and in the highest location of the southern population. Hydrological, Spring and Summer SPEI showed a strong positive correlation with tree-growth (Figure 7b), specially for the northern population (r > 0.6). Relationships with temperatures were lower than those with moisture variables. Yet there were differences between northern and southern populations: spring maximum temperature was the most significant limitation for the tree-growth of the southern population at both elevations while maximum temperatures of the previous August were the only negative significant for the northern one (Figures 7c and 7d). Remarkably minimum and maximum temperatures of current September positively influenced the tree-growth only in the northern population. No positive relationship with temperature was found for southern trees.

## Disturbances
The analysis of growth changes revealed differences between sites (Figures 8 and 9). En el sitio del norte osbervamos dos eventos de liberación importantes (GC > 50 %), que ocurren en mas de la mitad de los árboles muestreados (Figures 8 and 9). El primero de ellos desde 1940 hasta 1950, mientras que el segundo entre 1994 y 2001. Estos periodos se alternan con periodos de supresión. En los sitios del sur (caH y caL), sin embargo se observan algunos episodios débiles de supresión, aunque estos no se observan para los últimos años, de hecho incluso se observa una débil liberación (Figure 8) afectando a mas de la mitad de los árboles muestreados (Figure 9).
